---
title: "Advanced Time Series Models"
description: |
       Calibrate time-series forecasting models by using exponential smoothing and ARIMA techniques
  
author:
  - name: Raveena Chakrapani
    url: https://www.linkedin.com/in/raveena-chakrapani-444a60174/
    affiliation: School Of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objective

+ Import and wrangling time-series data by using appropriate tidyverse methods.

+ Visualise and analyse time-series data,calibrate time-series forecasting models by using exponential smoothing and ARIMAtechniques.

+ compare and evaluate the performance of forecasting models.

### Loading Packages

```{r}
packages = c('tidyverse', 'lubridate', 'zoo', 
               'seasonal', 'tsibble', 'feasts', 
               'fable','ggpubr', 'patchwork')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

### Importing Data

```{r}
tsbl_longer <- read_rds(
  "data/tsbl_longer.rds")
```

### Step 1: Time Series Data Sampling

```{r}
vietnam_ts <- tsbl_longer %>%
  filter(Country == "Vietnam") %>% 
  mutate(Type = if_else(
    `Month-Year` >= "2019-01-01", 
    "Hold-out", "Training"))
```

Next, a training data set is extracted from the original data set by using filter() of dplyr package.

```{r}
vietnam_train <- vietnam_ts %>%
  filter(`Month-Year` < "2019-01-01")
```

### Step 2: Exploratory Data Analysis (EDA): Time Series Data

```{r}
tsbl_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```
```{r}
tsbl_longer %>%
  filter(`Country` == "Australia") %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

Step 3: Fitting Exponential Smoothing State Space (ETS) Models: fable methods

In fable, Exponential Smoothing State Space Models are supported by ETS(). The combinations are specified through the formula:

```{r}
ETS(y ~ error(c("A", "M")) 
    + trend(c("N", "A", "Ad")) 
    + season(c("N", "A", "M")))
```
### Fitting a simple exponential smoothing (SES)

Adding the trend, seasonal pattern manually

```{r}
fit_ses <- vietnam_train %>%
  model(ETS(Arrivals ~ error("A") 
            + trend("N") 
            + season("N")))
fit_ses
```

### Examine Model Assumptions

```{r}
gg_tsresiduals(fit_ses)
```
### Model Details

```{r}
fit_ses %>%
  report()
```

### Fitting ETS Methods with Trend: Holt’s Linear

### Trend Methods

```{r}
vietnam_H <- vietnam_train %>%
  model(`Holt's method` = 
          ETS(Arrivals ~ error("A") +
                trend("A") + 
                season("N")))
vietnam_H %>% report()
```

### Dampened Trend Methods

```{r}
vietnam_HAd <- vietnam_train %>%
  model(`Holt's method` = 
          ETS(Arrivals ~ error("A") +
                trend("Ad") + 
                season("N")))
vietnam_HAd %>% report()
```

### Checking for Results

```{r}
gg_tsresiduals(vietnam_H)
```

```{r}
gg_tsresiduals(vietnam_HAd)
```